Structure constants and short-time simulations of open collective quantum spin systems

### 2019-04-04 soc_squeezing/collective_paper/collective_spin.tex
some wording changes in the collective spin paper

The time derivative operator $T$ will generally couple spin operators
$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, meaning we must somehow
truncate our equations of motion.  The simplest truncation strategy
would be to take
\begin{align}
  \f{d}{dt} \O\p{t}
  \to \sum_{w\p{\v m}<W} \S_{\v m}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
system of differential equations defined by Eq.~\eqref{eq:EOM}, and
allows us to solve it using standard numerical methods.  Some initial
conditions for this system of differential equations, i.e.~expectation
values of collective spin operators with respect to states that are
simple to prepare experimentally, are provided in Appendix
\ref{sec:initial_conditions}.

@@

Here compute the expectation values of all collective spin operators
with respect to spin-polarized (Gaussian) states.  These states are
parameterized by polar and azimuthal angles $\theta\in[0,\pi)$,
$\phi\in[0,2\pi)$, and lie within the Dicke manifold spanned by states
$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with
$S\equiv N/2$ and $S_\z\ket{k}=k\ket{k}$:
\begin{align}
  \ket{\theta,\phi}
  \equiv \sp{\cos\p{\theta/2} e^{i\phi/2} \ket\up
    + \sin\p{\theta/2} e^{-i\phi/2} \ket\dn}^{\otimes N}
  = \sum_{k=-S}^S { N \choose S+k }^{1/2}
  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{ik\phi} \ket{k}.
\end{align}
In order to compute expectation values of collective spin operators we
expand
\begin{align}
  S_\z = \sum_{k=-S}^S k \op{k},
  &&
  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
  g_\mu\p{k} \op{k+\mu}{k}
  = \sum_{k=-S+\delta_{\bmu,-1}}^{S-\delta_{\bmu,1}}
  g_\bmu\p{k} \op{k}{k+\bmu},
\end{align}
with $\bmu\equiv-\mu\in\set{+1,-1}$ and
\begin{align}
  g_\mu\p{k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
\end{align}
which implies
\begin{align}
  S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n
  &= \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
  \sp{\prod_{p=0}^{\ell-1} g_\mu\p{k+\mu p}}
  \sp{\prod_{q=0}^{n-1} g_\mu\p{k+\mu q}}
  \op{k+\mu\ell}{k+\mu n} \\
  &= \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
  \f{\p{S-\mu k}!}{\p{S+\mu k}!}
  \sp{\f{\p{S+\mu k+\ell}!}{\p{S-\mu k-\ell}!}
    \f{\p{S+\mu k+n}!}{\p{S-\mu k-n}!}}^{1/2}
  \op{k+\mu\ell}{k+\mu n} \\
  &= \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
  \f{\p{S-k}!}{\p{S+k}!}
  \sp{\f{\p{S+k+\ell}!}{\p{S-k-\ell}!}
    \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
  \op{\mu\p{k+\ell}}{\mu\p{k+n}}
\end{align}
We therefore have that
\begin{align}
  \bk{\theta,\phi|S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n|\theta,\phi}
  &= e^{i\phi \mu\p{\ell-n}} N! \sum_{k=-S}^{S-\max\set{\ell,n}}
  \f{k^m \p{S-k}! f_{\mu\ell n}\p{k,\theta}}
  {\p{S+k}! \p{S-k-\ell}! \p{S-k-n}!} \\
  &= e^{i\phi \mu\p{\ell-n}} \p{-1}^m N! \sum_{k=0}^{N-\max\set{\ell,n}}
  \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
  {k! \p{N-\ell-k}! \p{N-n-k}!}
\end{align}
for
\begin{align}
  f_{\mu\ell n}\p{k,\theta}
  \equiv \cos\p{\theta/2}^{N+\mu\p{2k+\ell+n}}
  \sin\p{\theta/2}^{N-\mu\p{2k+\ell+n}},
\end{align}
\begin{align}
  \tilde f_{\mu\ell n}\p{k,\theta}
  \equiv f_{\mu\ell n}\p{k-S,\theta}
  = \cos\p{\theta/2}^{2N\delta_{\mu,-1}+\mu\p{2k+\ell+n}}
  \sin\p{\theta/2}^{2N\delta_{\mu,1}-\mu\p{2k+\ell+n}}.
\end{align}
Defining the states
\begin{align}
  \ket{+\Z} \equiv \ket{0,0} = \ket\up^{\otimes N}, &&
  \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
  \ket{\X} \equiv \ket{\pi/2,0}
  = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
\end{align}
some particular expectation values of interest are
\begin{align}
  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
  = \delta_{\ell n} \times
  \begin{cases}
    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
    \delta_{n,0} \p{-S}^m & \mu \ne \nu,
  \end{cases},
\end{align}
and
\begin{align}
  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
  = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-\ell-k}!\p{N-n-k}!}.
\end{align}

>>>>

The time derivative operator $T$ will generally couple spin operators
$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, meaning we must somehow
truncate our equations of motion.  The simplest truncation strategy
would be to take
\begin{align}
  \f{d}{dt} \O\p{t}
  \to \sum_{w\p{\v m}<W} \S_{\v m}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
system of differential equations defined by Eq.~\eqref{eq:EOM}, and
allows us to solve it using standard numerical methods.  Some initial
conditions for this system of differential equations, namely
expectation values of collective spin operators with respect to
spin-polarized (Gaussian) states that are generally simple to prepare
experimentally, are provided in Appendix \ref{sec:initial_conditions}.

@@

Here we compute the expectation values of collective spin operators
with respect to spin-polarized (Gaussian) states.  These states are
parameterized by polar and azimuthal angles $\theta\in[0,\pi)$,
$\phi\in[0,2\pi)$, and lie within the Dicke manifold spanned by states
$\ket{k}\propto S_+^{S+k}\ket{\dn}^{\otimes N}$ with $S\equiv N/2$ and
$S_\z\ket{k}=k\ket{k}$:
\begin{align}
  \ket{\theta,\phi}
  \equiv \sp{\cos\p{\theta/2} e^{i\phi/2} \ket\up
    + \sin\p{\theta/2} e^{-i\phi/2} \ket\dn}^{\otimes N}
  = \sum_{k=-S}^S { N \choose S+k }^{1/2}
  \cos\p{\theta/2}^{S+k} \sin\p{\theta/2}^{S-k} e^{ik\phi} \ket{k}.
\end{align}
We can likewise expand, within the Dicke manifold,
\begin{align}
  S_\z = \sum_{k=-S}^S k \op{k},
  &&
  S_\mu = \sum_{k=-S+\delta_{\mu,-1}}^{S-\delta_{\mu,1}}
  g_\mu\p{k} \op{k+\mu}{k}
  = \sum_{k=-S+\delta_{\bmu,-1}}^{S-\delta_{\bmu,1}}
  g_\bmu\p{k} \op{k}{k+\bmu},
\end{align}
where $\bmu\equiv-\mu\in\set{+1,-1}$ and
\begin{align}
  g_\mu\p{k} \equiv \sqrt{\p{S-\mu k}\p{S+\mu k+1}},
\end{align}
which implies
\begin{align}
  S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n
  &= \sum_{k=-S+\delta_{\mu,-1}\max\set{\ell,n}}
  ^{S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
  \sp{\prod_{p=0}^{\ell-1} g_\mu\p{k+\mu p}}
  \sp{\prod_{q=0}^{n-1} g_\mu\p{k+\mu q}}
  \op{k+\mu\ell}{k+\mu n} \\
  &= \sum_{\mu k=-\mu S-\delta_{\mu,-1}\max\set{\ell,n}}
  ^{\mu S-\delta_{\mu,1}\max\set{\ell,n}} \p{\mu k}^m
  \f{\p{S-\mu k}!}{\p{S+\mu k}!}
  \sp{\f{\p{S+\mu k+\ell}!}{\p{S-\mu k-\ell}!}
    \f{\p{S+\mu k+n}!}{\p{S-\mu k-n}!}}^{1/2}
  \op{k+\mu\ell}{k+\mu n} \\
  &= \sum_{k=-S}^{S-\max\set{\ell,n}} k^m
  \f{\p{S-k}!}{\p{S+k}!}
  \sp{\f{\p{S+k+\ell}!}{\p{S-k-\ell}!}
    \f{\p{S+k+n}!}{\p{S-k-n}!}}^{1/2}
  \op{\mu\p{k+\ell}}{\mu\p{k+n}}.
\end{align}
We therefore have that
\begin{align}
  \bk{\theta,\phi|S_\mu^\ell \p{\mu S_\z^m} S_\bmu^n|\theta,\phi}
  &= e^{i\phi \mu\p{\ell-n}} N! \sum_{k=-S}^{S-\max\set{\ell,n}}
  \f{k^m \p{S-k}! f_{\mu\ell n}\p{k,\theta}}
  {\p{S+k}! \p{S-k-\ell}! \p{S-k-n}!} \\
  &= e^{i\phi \mu\p{\ell-n}} \p{-1}^m N! \sum_{k=0}^{N-\max\set{\ell,n}}
  \f{\p{S-k}^m \p{N-k}! \tilde f_{\mu\ell n}\p{k,\theta}}
  {k! \p{N-k-\ell}! \p{N-k-n}!}
\end{align}
with
\begin{align}
  f_{\mu\ell n}\p{k,\theta}
  \equiv \cos\p{\theta/2}^{N+\mu\p{2k+\ell+n}}
  \sin\p{\theta/2}^{N-\mu\p{2k+\ell+n}},
\end{align}
\begin{align}
  \tilde f_{\mu\ell n}\p{k,\theta}
  \equiv f_{\mu\ell n}\p{k-S,\theta}
  = \cos\p{\theta/2}^{2N\delta_{\mu,-1}+\mu\p{2k+\ell+n}}
  \sin\p{\theta/2}^{2N\delta_{\mu,1}-\mu\p{2k+\ell+n}}.
\end{align}
Defining the states
\begin{align}
  \ket{+\Z} \equiv \ket{0,0} = \ket\up^{\otimes N}, &&
  \ket{-\Z} \equiv \ket{\pi,0} = \ket\dn^{\otimes N}, &&
  \ket{\X} \equiv \ket{\pi/2,0}
  = \p{\f{\ket\up+\ket\dn}{\sqrt2}}^{\otimes N},
\end{align}
some particular expectation values of interest are
\begin{align}
  \bk{\nu\Z|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\nu\Z}
  = \delta_{\ell n} \times
  \begin{cases}
    \p{S-n}^m \f{N! n!}{\p{N-n}!} & \mu = \nu, \\
    \delta_{n,0} \p{-S}^m & \mu \ne \nu,
  \end{cases},
\end{align}
and
\begin{align}
  \bk{\X|S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n|\X}
  = \p{-1}^m \f{N!}{2^N} \sum_{k=0}^{N-\max\set{\ell,n}}
  \f{\p{S-k}^m \p{N-k}!}{k!\p{N-k-\ell}!\p{N-k-n}!}.
\end{align}
